---
title: "Code for Feasibility Document"
output: html_notebook
---

This document contains all of the code for my feasibility document and presentation. Figures created here are stored in the "Figs" folder within the repo. 

## Load in packages

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(plyr)
library(lubridate)
library(lme4)
library(Rmisc)
library(stats)
```

## Code for the date to day conversions: 

```{r}
# Internal process of the loop to convert days to dates
expt_day <- function(start.date, date){
  date <- mdy(date)
  start <- mdy(start.date)
  diff <- yday(date) - yday(start)

  return(diff)
}

# Code for the loop itself
loop <- function(start.date, date.list){
  days <- rep(NA, length(date.list))
  for (i in 1:(length(days))){
    days[i] <- expt_day(start.date = start.date, date = date.list[i])
  }
  return(days)
}
```


# Chapter 2: Success of C. congregata on different diet types

### Load in the Ch 2 data and clean it for the first figure

```{r}
ch2 <- read.csv("~/Desktop/GitHub/Feasibility/Data/MsCcSurvNov8.csv", header = TRUE)

ch2 <- ch2[ch2$survived == "yes", ]

# Remove failed parasitizations
ch2 <- ch2[ch2$ID != 34, ]
ch2 <- ch2[ch2$ID != 36, ]
ch2 <- ch2[ch2$ID != 102, ]
ch2 <- ch2[ch2$ID != 124, ]

# Get development times for wandering
ch2$day.hatch <- loop(start.date = "04/21/2020", date.list = ch2$date.hatch)
ch2$day.w <- loop(start.date = "04/21/2020", date.list = ch2$date.w)

ch2$delta_w <- ch2$day.w - ch2$day.hatch
ch2$mass_w <- ch2$mass.w
```

### Load in the Diamond data and clean it

```{r}
diamond <- read.csv("~/Desktop/GitHub/Feasibility/Data/Diamond AmNat Subset.csv")

diamond <- diamond[diamond$temp == 25, ]
diamond <- diamond[diamond$pop == "L", ]

diamond$food <- diamond$plant
diamond$ID <- diamond$cat_id

```

### Load in the diet data and clean it

```{r}
diet <- read.csv("~/Desktop/GitHub/Feasibility/Data/diet2006.csv")

diet$day.hatch <- loop(start.date = "07/26/2006", date.list = diet$date.hatch)
diet$day.w <- loop(start.date = "07/26/2006", date.list = diet$date.w)

diet$delta_w <- diet$day.w - diet$day.hatch
diet$mass_w <- diet$mass.w

```

## Figure 3

### 1. Create experiment column

```{r}
ch2$expt <- "Parker"
diamond$expt <- "Diamond"
diet$expt <- "Diamond"
```

### 2. Reduce to needed columns and combine

```{r}
a1 <- ch2 %>%
  select(expt, ID, food, delta_w, mass_w)

d1 <- diamond %>%
  select(expt, ID, food, delta_w, mass_w)

d2 <- diet %>%
  select(expt, ID, food, delta_w, mass_w)
```

```{r}
ch2g3 <- (rbind(a1, d1, d2))
```

### 3. Graph! 

```{r}
ggplot(aes(x = delta_w, y = mass_w, color = food, shape = expt), data = ch2g3) + 
  theme_bw() + 
  geom_point() + 
  scale_shape_manual(values = c(1, 4),  name = "Experiment") + 
  labs(x = "Time until wandering (days)", y = "Mass at wandering (mg)") + 
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Food Type",
                     label=c("Devil's Claw", "Diet", "Tobacco"))

ggsave("~/Desktop/GitHub/Feasibility/Figs/Ch2Fig3.png", width = 8, height = 5)
```

## Figure 4

### Clean the data for wasp analyses

```{r}
ch2g4 <- ch2 %>%
  select(ID, food, bin, num.para, num.w.coc, num.w.larvae, num.w.eclos.tot)

ch2g4 <- ch2g4[complete.cases(ch2g4), ]
```

### Make new columns for analyses

```{r}
ch2g4$num.w.emerg.tot <- ch2g4$num.w.coc + ch2g4$num.w.larvae
```

### Make means for each stage

```{r}
means <- ch2g4 %>%
  group_by(food) %>%
    summarise(emerged = mean(num.w.emerg.tot), cocoons = mean(num.w.coc), 
              eclosed = mean(num.w.eclos.tot))

meanslong <- gather(means, stage, number, emerged:eclosed)

meanslong$stage <- ordered(meanslong$stage, levels = c("emerged", "cocoons", "eclosed"))
```

```{r}
ch2g4$emerged <- ch2g4$num.w.emerg.tot 
ch2g4$cocoons <- ch2g4$num.w.coc 
ch2g4$eclosed <- ch2g4$num.w.eclos.tot

ch2g4 <- ch2g4 %>%
  select(ID, food, emerged, cocoons, eclosed)

ch2g4long <- gather(ch2g4, stage, number, emerged:eclosed)
ch2g4long$stage <- ordered(ch2g4long$stage, levels = c("emerged", "cocoons", "eclosed"))
```

```{r}
ses <- summarySE(ch2g4long, measurevar = "number",
                  groupvars = c("food", "stage"),
                  na.rm=TRUE)
```

### Graph!!

```{r}
ggplot(aes(x = stage, y = number, color = food), data = ses) + 
  theme_bw() + 
  geom_point(aes(size = 3)) + 
  geom_line(aes(group = food)) + 
  ylim(0, 120) + 
  geom_errorbar(aes(ymin = number - se, ymax = number + se),
                width = .1, size = .5) +
  scale_color_manual(values=c("#95D840", "#3F87BC", "#440D54"),
                     name="Food Type",
                     label=c("Devil's Claw", "Diet", "Tobacco")) + 
  guides(size = FALSE) + 
  labs(x = "Life stage", y = "Mean number of individuals")

ggsave("~/Desktop/GitHub/Feasibility/Figs/Ch2Fig4.png", width = 8, height = 5)
```

# Chapter 3: Leaf Consumption by different developmental outcomes

## Load in data

```{r}
ch3 <- read.csv("~/Desktop/GitHub/Feasibility/Data/MsCcHSFeedDec19final.csv")
scan <- read.csv("~/Desktop/GitHub/Feasibility/Data/MsCcHSFeedLeafScanFinal.csv", header = T)
```

## Manipulate data into correct forms

```{r}
ch3 <- ch3[ch3$fate != "dead" & ch3$fate != "quink" & ch3$fate != "wander*" & ch3$fate != "?", ]

ch3$outcome <- ch3$fate

ch3$outcome <- gsub(" mongo", "", ch3$outcome)
ch3$outcome <- gsub("em", "we", ch3$outcome)

```

### Get leaf area consumed
```{r}
#1
scan$comb <- paste(scan$Caterpillar.ID, "_", scan$Leaf.ID)

#0
unique <- unique(scan$comb)
consumed <- as.data.frame(unique)
consumed$area <- 0

#2-4
for (i in 1:length(consumed$unique)){
    subset <- filter(scan, scan$comb == consumed[i, 1])
    value <- max(subset$Leaf.Area) - min(subset$Leaf.Area)
    consumed[i, 2] <- value
}  

#5
consumed <- consumed %>%
  separate(unique, c("catID", "leafID"), "_")

#6
ID <- unique(consumed$catID)
consumedsum <- as.data.frame(ID)
consumedsum$totalarea <- 0

#7
for (i in 1:length(consumedsum$ID)){
  subset2 <- filter(consumed, consumed$catID == consumedsum[i, 1])
  value2 <- sum(subset2$area)
  consumedsum[i, 2] <- value2
}

#8
consumedsum$ID <- as.numeric(consumedsum$ID)

consumedsum <- consumedsum[consumedsum$ID %in% ch3$ID, ]

ch3 <- ch3[ch3$ID %in% consumedsum$ID, ]

consumedsum <- consumedsum[order(consumedsum$ID), ]

ch3$area <- consumedsum$totalarea

```

### Dates to days 

```{r}
ch3$day.hatch <- loop(start.date = "07/12/2020", date.list = ch3$date.hatch)
ch3$day.max <- loop(start.date = "07/12/2020", date.list = ch3$date.max)

ch3$ttmax <- ch3$day.max - ch3$day.hatch
```

### Wasp metrics

```{r}
ch3$perc.w.ecl <- ch3$num.w.eclos / (ch3$num.w.coc + ch3$num.w.larvae)
ch3$num.w.total <- (ch3$num.w.coc + ch3$num.w.larvae)

ch3$HS <- factor(ch3$HS, levels = c("yes", "no"))
```

## Figure 5 - Leaf area consumed by outcome

### From ASN presentation: 

ggplot(aes(x = area, y = mass.max), data = ch3) + 
  theme_bw() + 
  geom_point(aes(color = outcome, shape = food)) + 
  scale_shape_manual(values = c(19, 6)) + 
  scale_color_manual(values = c("#882255", "#88CCEE", "#117733")) +
  labs(x = "Leaf surface area consumed (cm2)", y = "Maximum mass achieved (mg)",
       color = "Developmental Outcome", shape = "Hostplant Consumed")

ggsave("~/Desktop/GitHub/Feasibility/Figs/Ch3Fig5.png", width = 8, height = 5)


### Alternate with matching colors: 

```{r}
ggplot(aes(x = area, y = mass.max), data = ch3) + 
  theme_bw() + 
  geom_point(aes(color = food, shape = outcome)) + 
  scale_shape_manual(values = c(6, 4, 19)) + 
  scale_color_manual(values = c("#95D840", "#440D54")) +
  labs(x = "Leaf surface area consumed (cm2)", y = "Maximum mass achieved (mg)",
       shape = "Developmental Outcome", color = "Hostplant Consumed")

ggsave("~/Desktop/GitHub/Feasibility/Figs/Ch3Fig5.png", width = 8, height = 5)
```


## Figure 6 - Caterpillar developmental data 


```{r}
ggplot(aes(x = ttmax, y = mass.max), data = ch3) + 
  theme_bw() + 
  geom_point(aes(color = food, shape = outcome)) + 
  scale_shape_manual(values = c(6, 4, 19)) + 
  scale_color_manual(values = c("#95D840", "#440D54")) +
  labs(x = "Time at maximum mass (days)", y = "Maximum mass achieved (mg)",
       shape = "Developmental Outcome", color = "Hostplant Consumed")

ggsave("~/Desktop/GitHub/Feasibility/Figs/Ch3Fig6.png", width = 8, height = 5)
```

## Figure 7 - Wasp survival to eclosion

```{r}
ggplot(aes(x = num.w.total, y = perc.w.ecl, shape = HS), data = ch3) + 
  theme_bw() + 
  ylim(0, 1.05) + 
  geom_smooth(method = "lm", se = F, aes(linetype = HS), color = "black") + 
  scale_linetype_manual(values = c("solid", "dotted")) + 
  geom_point(alpha = 0.8, aes(color = food)) + 
  scale_shape_manual(values = c(19, 4)) +
  scale_color_manual(values = c("#95D840", "#440D54")) + 
  labs(x = "Number emerged", y = "Percentage of wasps eclosed", shape = "Heat Shock Status", 
       color = "Hostplant Consumed", linetype = "Linear reg - Heat shock status")

ggsave("~/Desktop/GitHub/Feasibility/Figs/Ch3Fig7.png", width = 8, height = 5)
```
