---
title: "Statistics for my feasibility document"
output: html_notebook
---

Here, I will run statistics for my feasibility document. This will give me significances for my different calculations. 

## Load in packages

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(plyr)
library(lubridate)
library(lme4)
library(Rmisc)
library(stats)
```

## Code for the date to day conversions: 

```{r}
# Internal process of the loop to convert days to dates
expt_day <- function(start.date, date){
  date <- mdy(date)
  start <- mdy(start.date)
  diff <- yday(date) - yday(start)

  return(diff)
}

# Code for the loop itself
loop <- function(start.date, date.list){
  days <- rep(NA, length(date.list))
  for (i in 1:(length(days))){
    days[i] <- expt_day(start.date = start.date, date = date.list[i])
  }
  return(days)
}
```


# Chapter 2: Success of C. congregata on different diet types

## Load in the Ch 2 data and clean it

```{r}
ch2 <- read.csv("~/Desktop/GitHub/Feasibility/Data/MsCcSurvDec31Final.csv", header = TRUE)

ch2 <- ch2[ch2$survived == "yes", ]

# Remove failed parasitizations
ch2 <- ch2[ch2$ID != 34, ]
ch2 <- ch2[ch2$ID != 36, ]
ch2 <- ch2[ch2$ID != 102, ]
ch2 <- ch2[ch2$ID != 124, ]

# Get development times for wandering
ch2$day.hatch <- loop(start.date = "04/21/2020", date.list = ch2$date.hatch)
ch2$day.w <- loop(start.date = "04/21/2020", date.list = ch2$date.w)
ch2$day.coc <- loop(start.date = "04/21/2020", date.list = ch2$date.coc)

ch2$delta_w <- ch2$day.w - ch2$day.hatch
ch2$mass_w <- ch2$mass.w

ch2$delta_coc <- ch2$day.coc - ch2$day.hatch
ch2$mass_coc <- ch2$mass.frozen

ch2para <- ch2[ch2$para == "yes", ]
ch2unpara <- ch2[ch2$para == "no", ]
```

## Differences in mass and development time between treatment groups

### Parasitized mass at emergence

```{r}
mod2paramass <- lmer(mass_coc ~ food + (1|bin), data = ch2para)

summary(mod2paramass)
anova(mod2paramass)
```
Diet is lowest, then DC, then TB

Numerator df = 2
Denominator df = 55

F-value from table to be significant at alpha = 0.01 - ~5 

Groups are significantly different 

### Unparasitized mass at wandering

```{r}
mod2unparamass <- lmer(mass_w ~ food + (1|bin), data = ch2unpara)

summary(mod2unparamass)
anova(mod2unparamass)
```

DC is lowest, then diet, then TB

Not significantly different!

### Parasitized dev time at emergence

```{r}
mod2paradt <- lmer(day.coc ~ food + (1|bin), data = ch2para)

summary(mod2paradt)
anova(mod2paradt)
```

Overlapping

Fit is singular, but not significant. 

### Unparasitized dev time at wandering

```{r}
mod2unparadt <- lmer(day.w ~ food + (1|bin), data = ch2unpara)

summary(mod2unparadt)
anova(mod2unparadt)
```

TB, then diet, then DC

Not significant! 

## Survival of wasps between developmental stages

```{r}
ch2para$num.emerge.tot <- ch2para$num.w.coc + ch2para$num.w.larvae

ch2para$perc.w.surv.tot <- ch2para$num.w.eclos.tot / ch2para$num.emerge.tot
```

```{r}
mod2waspsurv <- glmer(perc.w.surv.tot ~ food + (1|bin), data = ch2para, family = "binomial", 
                      weights = num.emerge.tot )

summary(mod2waspsurv)
anova(mod2waspsurv)
```

TB significantly different than DC, diet is not  

ANOVA is very close to significance - need to bootstrap or something to figure it out 

### Percentages of wasps that had emergence at 4th instar in each diet type 

```{r}
ch2paracount <- ch2para %>%
  select(ID, food, mass.5th, mass_coc, num.w.coc, num.w.larvae, num.emerge.tot, num.w.eclos.tot, perc.w.surv.tot)

ch2paracount4 <- ch2paracount %>% 
  filter_all(any_vars(is.na(ch2paracount$mass.5th)))

ch2paracount5 <- ch2paracount %>%
  filter_all(any_vars(complete.cases(ch2paracount)))
```

```{r}
count(ch2paracount4[ch2paracount4$food == "diet", ])
# 9 
count(ch2paracount[ch2paracount$food == "diet", ])
#20

# 45%

count(ch2paracount4[ch2paracount4$food == "TB", ])
# 1
count(ch2paracount[ch2paracount$food == "TB", ])
#17

# 5.8%

count(ch2paracount4[ch2paracount4$food == "DC", ])
# 6 
count(ch2paracount[ch2paracount$food == "DC", ])
#19

# 31.5%
```

```{r}
t.test(ch2paracount4$num.emerge.tot, ch2paracount5$num.emerge.tot)
```

```{r}
t.test(ch2paracount4$mass_coc, ch2paracount5$mass_coc)
```

### Run model with diet type and para status predicting mass 

```{r}
ch2para$mass.final <- ch2para$mass.frozen

ch2unpara$mass.final <- ch2unpara$mass.w

ch2pmass <- ch2para %>% 
  select(ID, food, bin, para, mass.final)

ch2npmass <- ch2unpara %>%
  select(ID, food, bin, para, mass.final)

ch2mass <- rbind(ch2pmass, ch2npmass)
```

```{r}
mod2mass <- lmer(mass.final ~ food*para + (1|bin), data = ch2mass)

summary(mod2mass)
anova(mod2mass)
```

Yes, yes, yes - all significant 
